Characterization of the microbial and fungal communities in ancient pottery from Neolithic southwestern China

### 2024-02-10 analysis/paper/paper.qmd
work on relative abundance

## Taxonomic composition 
For the V4 region, bacterial composition at the phylum level in pottery interior samples was similar to those of pottery exterior samples and control samples, but slightly different from that of soil samples (@fig-V4-rel-ab). For Guijiabao pottery samples, *Firmicutes* (`r min(Firmi$rel_ab)`- `r max(Firmi$rel_ab)`%) was the most abundant phylum, followed by *Actinobacteria* (`r min(Actin$rel_ab)`- `r max(Actin$rel_ab)`%), *Deinococcus* (`r min(Deino$rel_ab)`- `r max(Deino$rel_ab)`%), and *Proteobacteria* (`r min(Prote$rel_ab)`- `r max(Prote$rel_ab)`%). Gaoshan sample is predominated by *Actinobacteria* (`r min(pot6_Actin$rel_ab)`- `r max(pot6_Actin$rel_ab)`%) with the presence of *Crenarchaeota* (`r min(pot6_Crena$rel_ab)`- `r max(pot6_Crena$rel_ab)`%) belonging to archaeal domain that was not found in Guijiabao samples and control samples. Although a different composition of phylum was observed between Guijiabao and Gaoshan samples, the difference is not statistically significant. The soil samples from Guijiabao is predominated by *Firmicutes* (`r min(Firmi_soil$rel_ab)`- `r max(Firmi_soil$rel_ab)`%) and *Thaumarchaeota* (`r min(Thau_soil$rel_ab)`- `r max(Thau_soil$rel_ab)`%) belonging to the archaeal domain, while soil samples from Gaoshan is predominated by *Actinobacteria* (`r Actin_soil4$rel_ab`- `r Actin_soil5$rel_ab`%) with less abundant *Thaumarchaeota* (`r Thau_soil4$rel_ab`- `r Thau_soil5$rel_ab`%). The dominant classes include *Bacilli*, *Clostridia* and *Negativicutes* among *Firmicutes*; *Actinobacteria*, *Rubrobacteria*, *Thermoleophilia* among *Actinobacteria*, *Deinococci* in *Deinococcus*, and *Gammaproteobacteria*, *Alphaproteobacteria*, and *Deltaproteobacteria* among *Proteobacteria*.

At the family level, significant differences were observed between pottery exterior samples and control samples, and between pottery exterior and interior samples. *Paenibacillaceae*, *Erysipelotrichaceae*, and *Xiphinematobacteraceae* were more abundant in pottery exterior samples than those in control samples (P <0.05). In addition to *Paenibacillaceae* and *Xiphinematobacteraceae*, *Phycisphaeraceae* was more abundant in pottery exterior samples than in pottery interior samples (P <0.05).

>>>>

## Taxonomic composition 
For the V4 region, bacterial composition at the phylum level in pottery interior samples was similar to those of pottery exterior samples and control samples, but slightly different from that of soil samples (@fig-V4-rel-ab). For Guijiabao pottery samples, *Firmicutes* (`r min(Firmi$rel_ab)`- `r max(Firmi$rel_ab)`%) was the most abundant phylum, followed by *Actinobacteria* (`r min(Actin$rel_ab)`- `r max(Actin$rel_ab)`%), *Deinococcus* (`r min(Deino$rel_ab)`- `r max(Deino$rel_ab)`%), and *Proteobacteria* (`r min(Prote$rel_ab)`- `r max(Prote$rel_ab)`%). Gaoshan sample is predominated by *Actinobacteria* (`r min(pot6_Actin$rel_ab)`- `r max(pot6_Actin$rel_ab)`%) with the presence of *Crenarchaeota* (`r min(pot6_Crena$rel_ab)`- `r max(pot6_Crena$rel_ab)`%) belonging to archaeal domain that was not found in Guijiabao samples and control samples. Although a different composition of phylum was observed between Guijiabao and Gaoshan samples, the difference is not statistically significant. The soil samples from Guijiabao is predominated by *Firmicutes* (`r min(Firmi_soil$rel_ab)`- `r max(Firmi_soil$rel_ab)`%) and *Thaumarchaeota* (`r min(Thau_soil$rel_ab)`- `r max(Thau_soil$rel_ab)`%) belonging to the archaeal domain, while soil samples from Gaoshan is predominated by *Actinobacteria* (`r Actin_soil4$rel_ab`- `r Actin_soil5$rel_ab`%) with less abundant *Thaumarchaeota* (`r Thau_soil4$rel_ab`- `r Thau_soil5$rel_ab`%).

The dominant classes include *Bacilli*, *Clostridia*, and *Negativicutes* among *Firmicutes*; *Actinobacteria*, *Rubrobacteria*, *Thermoleophilia* among *Actinobacteria*, *Deinococci* in *Deinococcus*, and *Gammaproteobacteria*, *Alphaproteobacteria*, and *Deltaproteobacteria* among *Proteobacteria*. At the family level, significant differences were observed between pottery exterior samples and control samples, and between pottery exterior and interior samples. *Paenibacillaceae*, *Erysipelotrichaceae*, and *Xiphinematobacteraceae* were more abundant in pottery exterior samples than those in control samples (P <0.05). In addition to *Paenibacillaceae* and *Xiphinematobacteraceae*, *Phycisphaeraceae* was more abundant in pottery exterior samples than in pottery interior samples (P <0.05). At the genus level, pottery exterior samples exhibit significantly higher abundance of *Candidatus_Xiphinematobacter* compared to pottery interior samples and control samples (P < 0.01). Among the bacterial taxa showing significant differences, only *Paenibacillaceae* displayed relative abundances exceeding 10% on average, while others were below 1%.
